\documentclass{letter}

\signature{Jason K. Moore}
\address{2420 37th St \\ Sacramento, CA 95817}

\begin{document}

\begin{letter}{
    Mech. and Aero. Engineering Faculty Search Committee \\
    University of California, Davis \\
    One Shields Avenue \\
  Davis, CA 95616}
\opening{Dear Search Committee:}

I am writing to apply for the Assistant Professor Faculty Position in
Prosthetics and Assistive Robotics. I am currently faculty in the same
department as the advertised position as a Lecturer with Potential Security of
Employment. I love teaching and the work I have done for the past three years,
but I long to have more time and resources for research. Three years ago I was
diverted from my post doctoral research work and my path towards a research
oriented faculty position and ended up back in the Sacramento region. Most of
my accomplishments since then have been teaching related, but I have managed to
keep up some research through small undergraduate projects and computational
based work with some of my prior collaborators. My post doctoral work focused
on developing data driven control algorithms for lower limb exoskeletons and
this position is an ideal fit to continue the momentum I started then. I have
included a research statement that shows how my work fits into the state of the
art research in powered prosthetics and assistive robotics along with a
teaching statement that focuses on reorientation of my teaching plans that
align with the position. Thank you for your consideration.

\closing{Sincerely,}

\end{letter}
\end{document}
