\documentclass{letter}

\date{September 17, 2018}
\signature{Jason K. Moore}
\address{2420 37th St \\ Sacramento, CA 95817}

\begin{document}

\begin{letter}{
    Mech. and Aero. Engineering Faculty Search Committee \\
    University of California, Davis \\
    One Shields Avenue \\
  Davis, CA 95616}
\opening{Dear Search Committee:}

I am writing to apply for the Assistant Professor Faculty Position in
Prosthetics and Assistive Robotics. I am currently faculty in the same
department as the advertised position as a Lecturer with Potential Security of
Employment.

My post doctoral work focused on developing data driven control algorithms for
lower limb exoskeletons and the advertised position is an ideal fit to continue
the momentum began then. Three years ago, I was diverted from that work and my
path towards a research oriented faculty position. I ended up back in the
Sacramento region and was fortunate to be hired as a teaching focused
professor. Most of my accomplishments since then have been teaching related,
but I have managed to keep up some research through small undergraduate
projects and computational based work with some of my prior collaborators. I
love teaching and the work I have done for the past three years, but I long to
have more time and resources for research.

I have included a research statement that shows how my work fits into the state
of the art research in powered prosthetics and assistive robotics and a
teaching statement that focuses on reorientation of my teaching plans that
align with the position. Thank you for your consideration.

\closing{Sincerely,}

\end{letter}
\end{document}
