\documentclass{article}

\usepackage[margin=1in]{geometry}
\usepackage[colorlinks=true]{hyperref}

\title{Teaching Statement}
\date{}

\begin{document}

%\maketitle

My interest in mechanical engineering began in my childhood where I spent many
hours making, tinkering, and discovering how the built environment functioned.
These hands-on experiences allowed me to develop a desire to discover the
connection to mathematics, physics, and science in order to improve my
projects. Once that realization occurred, any barriers I had towards learning
the engineering fundamentals were lifted. I hope to bring similar kinds of
experiences and motivation to the students I teach.

\section*{Teaching Philosophy}
%
Traditional engineering pedagogy focuses heavily on lecture-based theoretical
courses paired with few hands-on laboratory courses. The aspects of creativity,
team work, realistic constraints, and an infinite possibility of problem
solutions are all too often absent. These aspects are needed to develop the
practical side of engineering that students will require in their future
careers. Often, only the students who participate in extracurricular projects,
internships, or research positions make the connection from theory to practice
early on in the curriculum. Thus, the traditional structure may not be the best
way to maximize student understanding of engineering principles and instill the
agile problem solving methods they will need in the future. I strongly believe
that one cannot design if one cannot make, thus my philosophy aligns well with
Cal Poly's motto ``Learn by Doing''.

Although there are variety of motivational reasons students pursue engineering,
I believe introducing the applied practices of engineering early in education
will build passion and interest. Once this door is open, it is much easier to
weave in the skills the students need to become stronger engineers in their
future careers. I would like to see an engineering curriculum that mimics the
actual practice of engineering through an iterative pattern of posing realistic
problems followed by a search for the necessary fundamentals. This will
culminate in the application of new knowledge to arrive at a solution to the
problem. Like my childhood experiences, this model provides interesting
realistic problems that allow students to discover the engineering fundamentals
as opposed to presenting all of the fundamentals before the interesting
problems arrive.

Engineering students are capable of creating and solving problems when they
enter college. We should enrich their entire experience (especially in the first
year!) with challenges from real-world problems that leave them begging for the
knowledge and tools that they typically have to slog through during their first
years of school. Richard Miller, President of Olin College, often draws an
analogy between engineering students and violin students: ``Can you imagine not
playing the violin until your fourth year of study? Violinists start making
sounds with their instrument the first day of lessons.'' An engineering
curriculum could allow our students to draw the engineering bow across the
strings the minute they step into the classroom. For these early project-based
courses to be effective, however, the latest pedagogical developments must be
utilized to maximize learning potential.

\section*{Practical Classroom Examples}
%
In my courses, I try to provide students with open-ended problems that
lead into larger projects instead of problems designed for rote learning and
traditional exams. This approach more closely mimics the practice of
engineering. I combine this with rubric based assessments that set the bar for
mastery for improving student outcomes and effective assessment. I attempt to
have a good mixture of group and individual work, leaning more heavily toward
the former so students are prepared for the needs of industry. I also have been
working to orient my classrooms towards active learning. My best example is the
utilization of ``computational thinking'' that makes use of live coding in
class. I have setup a JupyterHub server that students log into via laptops,
tablets, and phones during class that provides an interactive engineering
computational environment. This allows access to my interactive textbook that
students use as a reading guide while I provide examples paired with short
computing exercises to periodically assess learning. I have developed a related
workshop for other practitioners with my colleague Allen Downey from Olin
College of Engineering which has been successful. Lastly, another very
important method that I make use of is rapid in-class assessment; at every
break, each student provides me with anonymous quick feedback: one line
comments that share what they didn't understand and what was effective. This
allows me to adjust my teaching after the break based on the feedback. I tie
this in with collected feedback before, during, and after the course to have
data to back my teaching decisions.

All of these methods are backed by evidence from education research. To keep
up-to-date on topics like these, I follow the education research literature,
especially the summary literature aimed at practicing educators and attend
``teach the teachers'' style workshops as much as possible. I have worked
closely with the UCD Center for Educational Effectiveness and the UCD
Engineering Education Learning Community these past four years to improve
student learning in my courses and for my department.

\section*{Prior Experience and Future Interests}
%
A teacher is often at their best when they know their material well. I spent
most of my graduate school years in the UCD MAE department and now four years
on the faculty making me intimately familiar with their undergraduate and
graduate curriculum. I have taught a number of the available courses as a
teaching assistant, lecturer, and professor. At the undergraduate level I have
strong experience with mechanics and machine design courses along with the
dynamics and controls curriculum. I have taught ``Engineering Graphics and
Design'', ``Manufacturing Processes'', Introduction to Mechanical Vibrations``,
''Mechanical Design``, ''Mechanical Systems Design Project``, ''Vehicle
Dynamics``, and ''Analysis, Simulation and Design of Mechatronic Systems``. I
have also taught ''Multibody Dynamics`` at the graduate level. I believe my
prior experience would enable me to teach all of the following Cal Poly ME
courses: 128, 129, 130, 211, 212, 228, 234, 236, 251, 305, 318, 326, 328, 329,
401, 402, 404, 405, 410, 416, 422, 423, 428, 429, 430, 441.

In my capacity as a teaching oriented professor I have worked on introducing
evidence based practices and innovations in the classroom. Some highlights from
the last four years are:

\begin{itemize}
    \setlength\itemsep{0.1em}
  \item developed a design competition and exchange program with Meijo
    University (Nagoya, Japan) on the cultural influences of robot and machine
    design
  \item flipped a mechanical vibrations class by utilizing ``computational
    thinking'' and project oriented learning with a custom designed interactive
    textbook and deployment through a JupyterHub server
  \item created a design studio classroom space that facilities active learning
    for our design courses
  \item created extensive rubric based assessment for written and oral
    communication in the capstone design course
  \item created a set of twenty Jupyter notebooks on multibody dynamics for
    in-class use and accompanying publicly available videos
  \item developed a transit bus bicycle rack design project which included
    reverse engineering, concept generation, and lightweight prototyping
  \item solicitation and mentoring of over 90 industry, government, and
    non-profit supported design projects spanning the mechanical engineering
    discipline
  \item co-awarded a \$5M Department of Education grant to create interactive
    OER engineering textbooks as part of the LibreTexts project
  \item co-wrote a book on teaching with Jupyter
  \item co-developed a workshop designed to teach STEM educators effective
    methods to teach with computation
\end{itemize}

There are at least three new undergraduate courses that I would like to
co-develop in the future that are influenced by my research endeavors: 1) a
first year problem solving with data, simulation, and engineering computation,
2) an upper level applied robotics and controls course, and 3) an upper level
elective focusing on project based prosthesis, vehicle, and human augmentation
design  Additionally, I would be very interested in teaching Cal Poly's bicycle
design course. I am one of the world's leading researchers and educators on
this topic and it would fit me nicely.

At the graduate level, I am also well prepared to teach many courses in
dynamics, control theory, biomechanics, optimization, software engineering, and
vehicle dynamics. At Cal Poly, my goal would be to bring courses in these
topics into the undergraduate elective possibilities. All of these courses can
be transformed into undergraduate accessible versions that will allow them to
experience cutting edge engineering. Some ideas would be to develop a course
focusing on the design, simulation, and optimization of legged biomechatronics
that aligned closely with my research. Students will learn about neuromuscular
modeling, mammalian gait, and get exposed to the latest tools in the field
(OpenSim, Biomechancal ToolKit, ROS/Gazebo, IPOPT, etc). Additionally, an
experimental biomechanics oriented course would also nicely complement the
computational oriented one to prepare students for applied work in the field
and the orthopedic biomechanics course already offered at Cal Poly.

My course topic strengths are not entirely based on UCD MAE offerings.
I have spent time at Delft University of Technology, Old Dominion University,
Cleveland State University, Stanford University, and with the Software
Carpentry non-profit where I have gleaned both new course ideas and
methodologies to provide stronger connections to industry. I have experience in
teaching computational methods for data science. I have given numerous
workshops and tutorials to scientists and engineers on simulation,
optimization, and data analysis. I have been trained by the Software Carpentry
organization in pedagogical methods and teach two-day workshops around the
world to introduce scientists and engineers to the best practices and methods
in scientific computing. The mechanical engineer of the future will be
additionally tasked with data driven engineering. The engineering curriculum
will need to adapt to bring data science into many of the core courses for our
students to stay competitive in the job market, which I am ready to do.

\end{document}
