\documentclass{letter}

\usepackage[margin=1in]{geometry}
\usepackage[colorlinks=true]{hyperref}

\date{}
\signature{Jason K. Moore}
\address{Jason K. Moore\\2420 37th St \\ Sacramento, CA USA 95817}

\begin{document}

\begin{letter}{
  Professor Rida T. Farouki \\
  Search Committee Chair \\
  Multibody \& System Dynamics -- Machine Design \& Robotics \\
  Department of Mechanical and Aerospace Engineering \\
  University of California, Davis \\
  One Shields Avenue \\
  Davis, CA 95616
}

\opening{Dear Search Committee:}

I am writing to apply for both positions in the call ``Two Assistant Professor
Positions in the areas of 1) Multibody \& System Dynamics and 2) Machine Design
\& Robotics'' in the Mechanical and Aerospace Engineering Department at the
University of California, Davis. I am currently an Assistant Professor of
Teaching Mechanical and Aerospace Engineering (LPSOE) in the same department.
My merit reviews and other external signs indicate I am excelling in my current
position, which is focused on advancing engineering education in our
department. I have been in this position for four academic years and my
accomplishments over that time align with this position's heavy teaching and
educational focus. I have spent 11 of my last 14 years at UC Davis and I have a
deep connection to the both the academic and regional communities. It is a
cherished place for me and I have vested interest in my career path here.

My primary goal at this stage in my academic career is to move into a typical
professorship where I can continue to excel at teaching and education but also
have the resources and support I need to focus on and further my research
program in human mobility started during my graduate and postdoctoral
positions. My current position has systematic hurdles that prevent me from
fully realizing my research potential even if I excel at my primary educational
duties. For example, many incoming graduate students want to work with me but
the lack of lab space is an immediate deterrent. This unfortunately leaves me
with a deep longing for the research I love. Regardless, I have been able to
maintain a steady research output with a mostly undergraduate team while also
excelling at my education related activities.

My prior very active research record and what I have maintained in the last
four years are reflected in the 10 peer reviewed journal publications, 16 peer
reviewed conference papers, 2 draft books, 12 software packages, 30+ invited
and conference talks, and over \$5M I have helped bring to the campus.  My
collective works have been cited over 550 times and I have an h-index of
12~\footnote{Google Scholar:
\href{http://tinyurl.com/jkm-gscholar}{tinyurl.com/jkm-gscholar}}, which is
favorably comparable to a number of faculty already in the department. I have
also demonstrated leadership in the dynamics, biomechanics, transportation
safety, and computing research communities through chairing conference
sessions, organizing conferences, editing journals, and acting on scientific
committees. With the resources these two faculty positions could offer I will
be able to launch from this already strong platform and develop as a key
research contributor to the department and campus.

The UCD MAE department is at a clear turning point in its life due to the
numerous retiring faculty in the past decade. These faculty defined what the
world knows us for. It is certainly time to bring in many new people with new
ideas but we also must steward our legacy strengths into the promising and
dynamic future of the next generation of mechanical and aerospace engineering.
Both advertised positions highlight this implicit desire. With this in mind, I
am in a very unique position due to my strong institutional knowledge of our
department and familiarity with the curriculum. I have collaborated with
numerous researchers at UC Davis and have a strong regional research and
engineering network tied to over a hundred projects I have been involved with
over the years. I have taught eight of our undergraduate courses and taken
thirteen of our courses while a graduate student. All of these courses are in
the areas of multibody \& system dynamics, machine design \& robotics, and
controls. My current course offerings are aligned precisely with the advertised
positions and are a natural fit.

I also bring in new ideas and external perspective. I have spent 8 years at
four other institutions. My postdoctoral training was from two of the leading
biomechanicists in the world (van den Bogert and Delp) and my Fulbright year
offered experience at TU Delft, one of the world's top technical universities.
These experiences have provided with me with a diverse academic background and
stimulus for innovation to help transform our curriculum and research
activities for the next 30 years.  For example, I plan to revive our lost
strength in biomechanics but born anew in the form of biomechatronics to lead
us into the future. Additionally, many of the retiring and retired professors
that these two positions are meant to fill roles for would vouch for my
likelihood of excellent stewardship of their legacy and the ability to infuse
modern research and engineering themes into the future curricula.

My current research trajectory is centered around developing data-driven
human-machine synergistic controllers for powered exoskeletons, powered
prostheses, personal mobility vehicles (electric bicycles, scooters,
wheelchairs, skateboards), and other assistive devices. These assistive
vehicles and devices will play a significant role in how everyone, particularly
the abled and disabled, young and old, get around both indoors and outdoors in
the future. I will weave these device and vehicle technology focused efforts
into the broader impact to transform our country's transportation system so
that we can have a clean, multi-modal, energy efficient, accessible modes to
move people from place to place indoors and outdoors. One aspect that this
position could enable is an in depth exploration of the role smart
self-balancing personal mobility vehicles and devices can play in enhancing the
active lifestyle of elderly individuals. Also, electric bike share system
companies have a growing interest in control augmented vehicles for safety and
the ability to autonomously return bicycles to charging stations. This is an
avenue that I plan to pursue aggressively with help from our Institute for
Transportation studies new BicyclingPlus research collaborative.

I have included my research plan that centers around human mobility and how it
fits here at UC Davis and a teaching statement that outlines my pedagogical
practices and how I plan to contribute to new curricula if hired.

Thank you for your time and consideration.

\closing{Sincerely,}

\end{letter}
\end{document}
