\documentclass{letter}

\usepackage[a4paper,margin=25mm]{geometry}

\date{}
\signature{Jason K. Moore}
\address{2420 37th St \\ Sacramento, CA USA 95817}

\begin{document}

\begin{letter}{
  Department of Biomechanical Engineering \\
  Delft University of Technology \\
  Delft, Netherlands
}

\opening{Dear Search Committee:}

I am writing to apply for the ``Assistant/Associate Professor (tenure track) in
Dynamics of Multibody systems'' in the Department of Biomechanical Engineering
at TU Delft. I believe that I am highly qualified for your department's
advertised position. For the last 15 years, I have developed unique research
expertise in the dynamics and control of bicycles as well as other topics that
align with your department's research areas. I have authored or co-authored 21
journal and conference articles on the subject of bicycles and my doctoral work
focused almost exclusively on the topic, resulting in a dissertation entitled
``Human Control of a Bicycle'' which is well cited. I have published 8 other
articles, many of which were ancillary to the bicycle dynamics and control
research. These works and my active role in the associated academic communities
make me one of the handful of people in the world with the expertise needed to
take the TU Delft Bicycle Dynamics and Control lab into the future.

I am currently an Assistant Professor of Teaching in the Mechanical and
Aerospace Engineering Department at the University of California, Davis. My
position is an academic senate tenure-track position primarily focused on
advancing undergraduate engineering education. I spend approximately 70\% of my
time on teaching and learning, 20\% on professional achievements and
activities, and 10\% on academic and public service. I have been in this
position for over four years and my accomplishments over those years align
mostly with this position's teaching focus but I have also maintained a
research program centered on human mobility in transportation, sports, and
assistive technologies with much of the work advancing engineering in the
dynamics and control of bicycles. Prior to this position, I was a postdoctoral
researcher in the field of human motion and control with applications in
powered lower limb exoskeletons. Before that, I was a graduate student with
research in bicycle dynamics, control and handling in a sport biomechanics lab.
Through these experiences I have developed strong applied computational and
experimental vehicle dynamics and biomechanics experience.

My current research trajectory is focused on developing human-machine
synergistic controllers for powered exoskeletons, powered prostheses, and
personal mobility vehicles, particularly single track vehicles. These assistive
devices will play a significant role in how abled and disabled individuals get
around in the future.

I have a strong record of obtaining competitive funding even though it is not a
requirement for my current position. As a graduate student, I initiated and
co-authored an awarded US National Science Foundation grant to study the
control of bicycles. This effort was unique given both the student initiative
and the fact that we were able to make the case for obtaining resources for a
research area that is often more challenging to fund. As a student, I was also
awarded a very competitive Fulbright Scholarship to the Netherlands, where I
spent a productive year in the bicycle dynamics lab at TU Delft. Most recently,
I was co-PI on a successful \$5 million grant from the US Department of
Education, as well as several internal grants for education activities over the
last four years. I plan to fund my future work through grants and industry
partnerships in the Netherlands, European Union, and internationally. I am
versed at the funding opportunities in the United States of America and am
looking forward to learning about the funding systems that support faculty
research in the Netherlands so that I can continue my successful trajectory.

My extensive teaching record in dynamics and design includes courses in the
topics of mechanical design, mechanical vibrations, system dynamics, multibody
dynamics, and vehicle dynamics at the undergraduate (BSc) and graduate (MSc,
PhD) levels. I make use of many education research backed practices in my
teaching and I am constantly improving the courses using the practices. I have
taught courses of up to 120 students and managed up to four teaching assistants
per class. I am also a leader in the use and promotion of computational
thinking for learning; recently co-authoring a book entitled ``Teaching and
Learning with Jupyter'' which provides and introduction to the related methods
and tools. I look forward to expanding this for the large courses at TU Delft.

I have also mentored just about 500 BSc and MSc students in over 100
engineering projects in the last four years. As these projects involved
external project sponsors I developed relationships with a diverse set of
research, non-profit, and industry organizations. Several of these projects are
also with international groups in Kenya, Nicaragua, Cambodia, and Sweden. I am
especially proud of the mechanical design exchange program I have developed
with Meijo University in Japan. As my native language is English, I will be
able to teach MSc classes immediately at TU Delft, but I also have a beginner
understanding of Dutch and am eager to become fluent so I can teach BSc
courses.

Lastly, I have a strong service record. I serve on my department's
undergraduate curriculum committee where we are modernizing our course
offerings. I serve on the scientific and planning committees of the Bicycle and
Motorcycle Dynamics conference series and recently hosted the International
Cycling Safety Conference at UC Davis for the first offering outside of Europe.
On the education front, I am a topic editor of the innovative Journal of Open
Source Education which is now in its second year of publishing.

The best career fit for me is to be a faculty member at a university department
that is mission focused, works together forwarding this common mission, and
strives for continuous change all while deeply valuing engineering education
through teaching and research mentorship.

If at TU Delft, I see numerous opportunities for collaboration within the
university. And frankly, it is exciting that so many authors of papers I have
read are at the institution. In the Biomechanical Engineering department, Prof.
van der Helm and Prof. Harlaar's expertise in upper and lower limb biomechanics
will provide insight for bicycling biomechanics. Prof. Vallery's work in
biomechatronics and teaching dynamics can align nicely with this position's
efforts. I have worked with Prof. Seth when he was at Stanford and his
foundational work and expertise in musculoskeletal simulation are strong
interests of mine. I have studied and cited Prof. van der Kooij's control
identification work and would be excited to learn directly from him.  Dr.
Geijtenbeek's predictive simulation methods and tools are leading this research
area and  his connections to Motek Medical could be helpful for my experimental
work. I have co-authored a paper with Prof. Happee in the Department of
Transport and Planning and will seek collaborations related to vehicle
engineering. The Control and Simulation group has strong expertise in manual
control theory and identification that I have leveraged in the past. For
example, I work some with Prof. van Paassen on the Python control software and
could expand that effort and Prof. Mulder's work in manual control has
benefited my applications to single track vehicles. Lastly, I am excited to
develop collaborations with Dutch bicycle companies being that the country
leads the world in innovative utility bicycle designs.

I have included my research plan that centers around human mobility and a
teaching statement that outlines my pedagogical practices in addition to how I
think I would fit into your department.

Thank you for your consideration.

\closing{Sincerely,}

\end{letter}
\end{document}

%- [x] internationally oriented
%- [x] successes in competitve funding
%- [x] international partnerships: academic and industry
%- [x] strengthen their existing reserach lines and add expertise
%- produce PhDs and [x] pubs
%- [x] teach bs and ms dynamics courses
%- [x] supervise bs, Mmc projects and [ ] supervise phds
%- [x] innovate teaching for large classses
%- [x]] service to scientific communities society, univseristy
%- [x] work with non profits
%- promote tu delft
%- [x] experience running lab doing experiments using modern equip
%- [x] exp teaching BSc and msc
%- like to create and innovate in university context
%- good communicator: enthusiam empahty, comminit passion
%- you can inspiree and motivate, coach and delegate, deal with diversity
%- [x] english
%- [x] willingnesss to leanr dutch

