\documentclass{letter}

\usepackage[margin=1in]{geometry}

\date{}
\signature{Jason K. Moore}
\address{2420 37th St \\ Sacramento, CA USA 95817}

\begin{document}

\begin{letter}{
  School \\
  Address
}

\opening{Dear Search Committee:}

I am writing to apply for a ``Tenure Track Faculty Position in Mechanical and
Aerospace Engineering'' in your department. I am currently faculty in the
Mechanical and Aerospace Engineering Department at the University of
California, Davis. My official title is ``Lecturer with Potential Security of
Employment'' which is an academic senate tenure-track position primarily
focused on advancing undergraduate engineering education. I have been in this
position for almost four years and my recent accomplishments align mostly with
this position's teaching needs, which are detailed in my last merit review
candidate statement (also attached).

My goal is to move into a more typical professorship where I can continue to
excel at teaching and education but also have the resources I need to further
my research in human mobility started during my graduate and postdoctoral
positions. The best fit for me is at a university has a department who is
mission focused, works together forwarding this common mission, and values
continuous change all while deeply valuing engineering education through
teaching and research mentorship.

My current research trajectory is focused on developing human-machine
synergistic controllers for powered exoskeletons, powered prostheses, and
human-scale mobility vehicles, particularly single track vehicles. These
assistive devices will play a significant role in how the abled and disabled
get around in the future.

I have included my research plan that centers around human mobility and a
teaching statement that outlines my pedagogical practices in addition to how I
think I would fit into your department.

Thank you for your consideration.

\closing{Sincerely,}

\end{letter}
\end{document}
