\documentclass{letter}

\usepackage[margin=1in]{geometry}

\date{}
\signature{Jason K. Moore}
\address{2420 37th St \\ Sacramento, CA 95817}

\begin{document}

\begin{letter}{
  Mech. and Aero. Engineering Faculty Search Committee \\
  University of California, Davis \\
  One Shields Avenue \\
  Davis, CA 95616}
\opening{Dear Search Committee:}

I am writing to apply to the two Assistant Professor Faculty Positions in
Mechatronics and System Dynamics and Controls. I am currently faculty in the
Mechanical and Aerospace Engineering Department at the University of
California, Davis. My official title is ``Lecturer with Potential Security of
Employment'' which is an academic senate tenure-track position primarily
focused on undergraduate engineering education.

I understand that this application is quite late in your process but after
meeting the Cal Poly Bicycle Builders club at the North American Handbuilt
Bicycle show in Sacramento this past month I was inspired to apply anyways.

With bicycles in mind, I am well aware of Bill Patterson's Cal Poly course on
single track vehicle design, which I have studied in detail. My PhD
dissertation was on human control of bicycles. I see that the course is still
being taught, which is a unique find in mechanical engineering departments. I
have fantasized about creating a similar course over the years.

The best fit for me is a university or college that deeply values engineering
education and that has a department who is mission focused, works together, and
values continuous change. I deeply believe in the university's motto ``Learn by
Doing'', as I hope you can see through my work.

I have included a teaching statement that outlines my pedagogical practices and
how I think I would fit into your department. I have taught many similar
courses to those in your offerings and have some new offering ideas that I hope
you find as attractive possible additions. Additionally, I have included my
research interests that centers around human mobility which I primarily support
undergraduate researchers in that work.

Thank you for your consideration.

\closing{Sincerely,}

\end{letter}
\end{document}
